---
title: "Alps Snowfall"
author: "Van Thuan Romoli"
date: "2024-08-11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#PACKAGES LOADING
library(ggplot2)
library(ggmap)
library(texmex)
library(ismev)
```

# Loading and visualizing the data

Our data is a large list consisting of variables such as air temperature, precipitations, etc. Each of these variables is organised as a 49x29x174 matrix where 49x29 is a spatial grid and 174 indicates the month.

Below, we extract the data for snow fall.

```{r}
load("Alps_precip.Rdata")
load("elev.Rdata")
image(data$snow_fall[,,1])

df <- data.frame("snowfall"=c(data$snow_fall[,,1]),
                 "lon"=c(lonlat[,,1]),
                 "lat"=c(lonlat[,,2])
)

indbox = make_bbox(lon = c(min(df$lon), max(df$lon))+c(-3,3), lat = c(max(df$lat), min(df$lat)), f = 0)
world <- map_data("world")
p = ggplot() +
  geom_map(
    data = world, map = world,
    aes(long, lat, map_id = region),
    color = "black", fill = "lightgray", size = 0.1
  ) + scale_x_continuous(limits=indbox[c(1,3)]) + scale_y_continuous(limits=indbox[c(2,4)]) + coord_fixed(ratio=1.4) +
geom_tile(df, mapping = aes(x = lon, y = lat, fill = snowfall), width = 0.27, height = 0.27) 
p


```

To be able to analyse this data set, we need to reshape from a grid to a long data frame, where each column contains all values for a variable.

```{r}
grid_to_vector <- function(mydataset) {
  h = 1
  temp_list <- vector(length = length(mydataset))
  for (i in 1:dim(mydataset)[1]) {
    for (j in 1:dim(mydataset)[2]) {
       for (k in 1:dim(mydataset)[3]) {
         temp_list[h] <- mydataset[i,j,k]
         h = h + 1
      }
    }
  }
  return(temp_list)
}
```

```{r}
snow_fall <- grid_to_vector(data$snow_fall)
air_temp <- grid_to_vector(data$air_temp)
precip <- grid_to_vector(data$precip)
wind_u <- grid_to_vector(data$wind_u)
wind_v <- grid_to_vector(data$wind_v)
dewpoint_temp <- grid_to_vector(data$dewpoint_temp)
anisotropy <- grid_to_vector(data$anisotropy)
slope <- grid_to_vector(data$slope)
sd <- grid_to_vector(data$sd)
isotropy <- grid_to_vector(data$isotropy)
land_sea <- grid_to_vector(data$`land-sea`)
mean_sealevelpressure <- grid_to_vector(data$mean_sealevelpressue)
geopotential <- grid_to_vector(data$geopotential)

lon <- vector(length = 49 * 29 * 174)
h = 1
for (i in 1:49) {
  for (j in 1:29) {
    for  (k in 1:174){
      lon[h] <- lonlat[i,j,1]
      h = h+1
    }
  }
}

lat <- vector(length = 49 * 29 * 174)
h = 1
for (i in 1:49) {
  for (j in 1:29) {
    for  (k in 1:174){
      lat[h] <- lonlat[i,j,2]
      h = h+1
    }
  }
}

alt <- vector(length = 49 * 29 * 174)
h = 1
for (i in 1:49) {
  for (j in 1:29) {
    for  (k in 1:174){
      alt[h] <- elev[i,j]
      h = h+1
    }
  }
}

months <- c(rep(1:12,14),c(1,2,3,4,5,6))
```

```{r}
alps_df <- data.frame(air_temp,alt,wind_u,wind_v,dewpoint_temp,anisotropy,slope,sd,isotropy,land_sea,mean_sealevelpressure,geopotential,dates,months,lon,lat,snow_fall,precip)
alps_df <- na.omit(alps_df)
```

# Fitting a Generalized Pareto Distribution

## Normalising the data set and choosing threshold

```{r}
alps_extremes <- alps_df[alps_df$snow_fall > 0,]
alps_extremes$snow_fall <- alps_extremes$snow_fall * 1000
 
# Normalising data set
for (i in 1:12) {
  alps_extremes[,i] <- (alps_extremes[,i] - mean(alps_extremes[,i])) / sd(alps_extremes[,i])
}


hist(alps_extremes$snow_fall)

mrl.plot(alps_extremes$snow_fall)

c(quantile(alps_extremes$snow_fall,0.5),quantile(alps_extremes$snow_fall,0.99))

gpd.fitrange(alps_extremes$snow_fall,0.4254007,6.2956244)

u <- 6
```

## Intercept-only model

```{r}
# Intercept model only
fit0 <- gpd.fit(xdat = alps_extremes$snow_fall, threshold = u)
gpd.diag(fit0)

print(2 * fit0$nllh + 2 * 2)

```

We now want to do model selection. We use forward step-wise selection and use two different methods to decide which covariates to add to the model.

## Model selection using AIC

### One covariate

```{r}

fit1covariate <- list()
for (i in 1:12) {
  fit1covariate[[i]] <- gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes[,i]),
                threshold = u, 
                sigl = c(1), siglink = exp)
}

for (i in 1:12) {
  print(c(2 * fit1covariate[[i]]$nllh + 2 * 3, names(alps_extremes)[i]))
}

# Add second covariate, i.e. ydat = altitude

gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt),
                threshold = u, 
                sigl = c(1), siglink = exp)


gpd.diag(fit1covariate[[2]])

print(fit1covariate[[2]]$nllh * 2 + 2 * 3)
```

### Two covariates

```{r}

fit2covariate <- list()

for (i in 1:12) {
  if (colnames(alps_extremes)[i] != "alt"){
    fit2covariate[[i]] <- gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt,
                             alps_extremes[,i]),
                threshold = u, 
                sigl = c(1,2), siglink = exp)
  }
}


for (i in 1:12) {
  print(c(2 * fit2covariate[[i]]$nllh + 2 * 4,names(alps_extremes)[i]))
}

# Add ninth covariate, i.e. ydat = altitude + isotropy

gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt,
                             alps_extremes$isotropy),
                threshold = u, 
                sigl = c(1,2), siglink = exp)

gpd.diag(fit2covariate[[9]])

print(fit2covariate[[9]]$nllh * 2 + 2 * 4)
#cor(alps_extremes$alt,alps_extremes$anisotropy, method = "spearman")

```

### Three covariates

```{r}

fit3covariate <- list()

for (i in 1:12) {
  if (colnames(alps_extremes)[i] != "alt" &
      colnames(alps_extremes)[i] != "isotropy"){
    fit3covariate[[i]] <- gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt,
                             alps_extremes$isotropy,
                             alps_extremes[,i]),
                threshold = u, 
                sigl = c(1,2,3), siglink = exp)
  }
}


for (i in 1:12) {
  print(c(2 * fit3covariate[[i]]$nllh + 2 * 5,names(alps_extremes)[i]))
}

# Add tenth covariate, i.e. ydat = altitude + isotropy + landsea

gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt,
                             alps_extremes$isotropy,
                             alps_extremes$land_sea),
                threshold = u, 
                sigl = c(1,2,3), siglink = exp)

gpd.diag(fit3covariate[[10]])

print(fit3covariate[[10]]$nllh * 2 + 2 * 5)
```

### Four covariates

```{r}
# Step-wise regression
fit4covariate <- list()

for (i in 1:12) {
  if (colnames(alps_extremes)[i] != "alt" &
      colnames(alps_extremes)[i] != "isotropy" &
      colnames(alps_extremes)[i] != "land_sea"){
    fit4covariate[[i]] <- gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt,
                             alps_extremes$isotropy,
                             alps_extremes$land_sea,
                             alps_extremes[,i]),
                threshold = u, 
                sigl = c(1,2,3,4), siglink = exp)
  }
}


for (i in 1:12) {
  print(c(2 * fit4covariate[[i]]$nllh + 2 * 6,names(alps_extremes)[i]))
}

# Add fifth covariate, i.e. ydat = altitude + isotropy + land sea + dewpoint temp

gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt,
                             alps_extremes$isotropy,
                             alps_extremes$land_sea,
                             alps_extremes$dewpoint_temp),
                threshold = u, 
                sigl = c(1,2,3,4), siglink = exp)

gpd.diag(fit4covariate[[5]])

print(fit4covariate[[5]]$nllh * 2 + 2 * 6)
```

### Five covariates

```{r}

fit5covariate <- list()

for (i in 1:12) {
  if (colnames(alps_extremes)[i] != "alt" &
      colnames(alps_extremes)[i] != "isotropy" &
      colnames(alps_extremes)[i] != "land_sea" &
      colnames(alps_extremes)[i] != "dewpoint_temp"){
    fit5covariate[[i]] <- gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt,
                             alps_extremes$isotropy,
                             alps_extremes$land_sea,
                             alps_extremes$dewpoint_temp,
                             alps_extremes[,i]),
                threshold = u, 
                sigl = c(1,2,3,4,5), siglink = exp)
  }
}


for (i in 1:12) {
  print(c(2 * fit5covariate[[i]]$nllh + 2 * 7,names(alps_extremes)[i]))
}

# Add first covariate, i.e. ydat = altitude + isotropy + land sea + dewpoint temp + air temp

gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt,
                             alps_extremes$isotropy,
                             alps_extremes$land_sea,
                             alps_extremes$dewpoint_temp,
                             alps_extremes$air_temp),
                threshold = u, 
                sigl = c(1,2,3,4,5), siglink = exp)

gpd.diag(fit5covariate[[1]])

print(fit5covariate[[1]]$nllh * 2 + 2 * 7)
```

### Six covariates

```{r}
# Step-wise regression
fit6covariate <- list()

for (i in 1:12) {
  if (colnames(alps_extremes)[i] != "alt" &
      colnames(alps_extremes)[i] != "isotropy" &
      colnames(alps_extremes)[i] != "land_sea" &
      colnames(alps_extremes)[i] != "dewpoint_temp" &
      colnames(alps_extremes)[i] != "air_temp"){
    fit6covariate[[i]] <- gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt,
                             alps_extremes$isotropy,
                             alps_extremes$land_sea,
                             alps_extremes$dewpoint_temp,
                             alps_extremes$air_temp,
                             alps_extremes[,i]),
                threshold = u, 
                sigl = c(1,2,3,4,5,6), siglink = exp)
  }
}


for (i in 1:12) {
  print(c(2 * fit6covariate[[i]]$nllh + 2 * 8,names(alps_extremes)[i]))
}

# Add eighth covariate, i.e. ydat = altitude + isotropy + land sea + dewpoint temp + air temp + sd

gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt,
                             alps_extremes$isotropy,
                             alps_extremes$land_sea,
                             alps_extremes$dewpoint_temp,
                             alps_extremes$air_temp,
                             alps_extremes$sd),
                threshold = u, 
                sigl = c(1,2,3,4,5,6), siglink = exp)

gpd.diag(fit6covariate[[8]])

print(fit6covariate[[8]]$nllh * 2 + 2 * 8)
```

### Seven covariates

```{r}
# Step-wise regression
fit7covariate <- list()

for (i in 1:12) {
  if (colnames(alps_extremes)[i] != "alt" &
      colnames(alps_extremes)[i] != "isotropy" &
      colnames(alps_extremes)[i] != "land_sea" &
      colnames(alps_extremes)[i] != "dewpoint_temp" &
      colnames(alps_extremes)[i] != "air_temp" &
      colnames(alps_extremes)[i] != "sd"){
    fit7covariate[[i]] <- gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt,
                             alps_extremes$isotropy,
                             alps_extremes$land_sea,
                             alps_extremes$dewpoint_temp,
                             alps_extremes$air_temp,
                             alps_extremes$sd,
                             alps_extremes[,i]),
                threshold = u, 
                sigl = c(1,2,3,4,5,6,7), siglink = exp)
  }
}


for (i in 1:12) {
  print(c(2 * fit7covariate[[i]]$nllh + 2 * 9,names(alps_extremes)[i]))
}

# Add seventh covariate, i.e. ydat = altitude + isotropy + land sea + dewpoint temp + air temp + sd + slope

gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt,
                             alps_extremes$isotropy,
                             alps_extremes$land_sea,
                             alps_extremes$dewpoint_temp,
                             alps_extremes$air_temp,
                             alps_extremes$sd,
                             alps_extremes$slope),
                threshold = u, 
                sigl = c(1,2,3,4,5,6,7), siglink = exp)

gpd.diag(fit7covariate[[7]])

print(fit7covariate[[7]]$nllh * 2 + 2 * 9)
```

### Eight covariates

```{r}
# Step-wise regression
fit8covariate <- list()

for (i in 1:12) {
  if (colnames(alps_extremes)[i] != "alt" &
      colnames(alps_extremes)[i] != "isotropy" &
      colnames(alps_extremes)[i] != "land_sea" &
      colnames(alps_extremes)[i] != "dewpoint_temp" &
      colnames(alps_extremes)[i] != "air_temp" &
      colnames(alps_extremes)[i] != "sd" &
      colnames(alps_extremes)[i] != "slope"){
    fit8covariate[[i]] <- gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt,
                             alps_extremes$isotropy,
                             alps_extremes$land_sea,
                             alps_extremes$dewpoint_temp,
                             alps_extremes$air_temp,
                             alps_extremes$sd,
                             alps_extremes$slope,
                             alps_extremes[,i]),
                threshold = u, 
                sigl = c(1,2,3,4,5,6,7,8), siglink = exp)
  }
}


for (i in 1:12) {
  print(c(2 * fit8covariate[[i]]$nllh + 2 * 10,names(alps_extremes)[i]))
}

# Add twelveth covariate, i.e. ydat = altitude + isotropy + land sea + dewpoint temp + air temp + sd + slope + geopotential

gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt,
                             alps_extremes$isotropy,
                             alps_extremes$land_sea,
                             alps_extremes$dewpoint_temp,
                             alps_extremes$air_temp,
                             alps_extremes$sd,
                             alps_extremes$slope,
                             alps_extremes$geopotential),
                threshold = u, 
                sigl = c(1,2,3,4,5,6,7,8), siglink = exp)

gpd.diag(fit8covariate[[12]])

print(fit8covariate[[12]]$nllh * 2 + 2 * 10)
```

### Nine covariates

```{r}
# Step-wise regression
fit9covariate <- list()

for (i in 1:12) {
  if (colnames(alps_extremes)[i] != "alt" &
      colnames(alps_extremes)[i] != "isotropy" &
      colnames(alps_extremes)[i] != "land_sea" &
      colnames(alps_extremes)[i] != "dewpoint_temp" &
      colnames(alps_extremes)[i] != "air_temp" &
      colnames(alps_extremes)[i] != "sd" &
      colnames(alps_extremes)[i] != "slope" &
      colnames(alps_extremes)[i] != "geopotential"){
    fit9covariate[[i]] <- gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt,
                             alps_extremes$isotropy,
                             alps_extremes$land_sea,
                             alps_extremes$dewpoint_temp,
                             alps_extremes$air_temp,
                             alps_extremes$sd,
                             alps_extremes$slope,
                             alps_extremes$geopotential,
                             alps_extremes[,i]),
                threshold = u, 
                sigl = c(1,2,3,4,5,6,7,8,9), siglink = exp)
  }
}


for (i in 1:11) {
  print(c(2 * fit9covariate[[i]]$nllh + 2 * 11,names(alps_extremes)[i]))
}

# No more adding since there isn't an improvement in the AIC
```

### Trying interaction terms

```{r}
# Step-wise regression
fit.int1.covariate <- list()
z = 1
for (i in 1:12) {
  for (j in 1:12) {
    fit.int1.covariate[[z]] <- gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt,
                             alps_extremes$isotropy,
                             alps_extremes$land_sea,
                             alps_extremes$dewpoint_temp,
                             alps_extremes$air_temp,
                             alps_extremes$sd,
                             alps_extremes$slope,
                             alps_extremes$geopotential,
                             alps_extremes[,i]*alps_extremes[,j]),
                threshold = u, 
                sigl = c(1,2,3,4,5,6,7,8,9), siglink = exp)
    z = z+1
  }
}

nllh <- numeric()
z = 1
for (i in 1:12) {
  for (j in 1:12) {
    print(c(2 * fit.int1.covariate[[z]]$nllh + 2 * 11,names(alps_extremes)[i],names(alps_extremes)[j]))
    nllh[z] <- 2 * fit.int1.covariate[[z]]$nllh + 2 * 11
    z = z + 1

  }
}


min(nllh)

# ydat = altitude + isotropy + land sea + dewpoint temp + air temp + sd + slope + geopotential + land_sea*sd

int1 <- gpd.fit(xdat = alps_extremes$snow_fall,
                ydat = cbind(alps_extremes$alt,
                             alps_extremes$isotropy,
                             alps_extremes$land_sea,
                             alps_extremes$dewpoint_temp,
                             alps_extremes$air_temp,
                             alps_extremes$sd,
                             alps_extremes$slope,
                             alps_extremes$geopotential,
                             alps_extremes$land_sea*alps_extremes$sd),
                threshold = u, 
                sigl = c(1,2,3,4,5,6,7,8,9), siglink = exp)

gpd.diag(int1)

print(int1$nllh * 2 + 2 * 11)
```

More interaction terms could be added; however, this adds complexity to the model. We chose to stick with a simpler model, with only one interaction term.

## k-fold cross-validation

```{r}
shuffleshuffle <- function() {
  
  tobereturned <- list()
  
  # Removes temporal dependencies and thresholds the data
  alps_shuffled <- alps_extremes[sample(1:nrow(alps_extremes), replace = FALSE),]
  alps_shuffled <- alps_shuffled[alps_shuffled$snow_fall > 6,]
  
  # Separates dataframe in 5 folds
  folds <- list()
  for (i in 1:5) {
    folds[[i]] <- alps_shuffled[((nrow(alps_shuffled)*((i-1)/5))+1):(nrow(alps_shuffled)*(i/5)),]
  }
  # Groups together 4 of the 5 folds for fitting and leaves out 1 for evaluation
  foldtraining <- list()
  for (i in 1:5) {
    foldtraining[[i]] <- rbind(folds[[(((i) %% 5) + 1)]],folds[[(((i+1) %% 5) + 1)]],folds[[(((i+2) %% 5) + 1)]],folds[[(((i+3) %% 5) + 1)]])
  
    #print(c(  (((i) %% 5) + 1), (((i+1) %% 5) + 1), (((i+2) %% 5) + 1), (((i+3) %% 5) + 1)                 ))
  }
  
  tobereturned[[1]] <- folds
  tobereturned[[2]] <- foldtraining
  
  return(tobereturned)
}
```

### One covariate

```{r}
mylistpacked <- list()
for (k in 1:50) {
  fit1.5kfold <- list()
  folds <- shuffleshuffle()[[1]]
  foldtraining <- shuffleshuffle()[[2]]
  for (j in 1:12) {
    nllh1.5kfold <- numeric()
    fit <- list()
    for (i in 1:5) {
      fit[[i]] <- gpd.fit(xdat = foldtraining[[i]]$snow_fall,
                          ydat = cbind(foldtraining[[i]][,j]),
                          threshold = 6, 
                          sigl = c(1), siglink = exp)
      X_1 <- cbind(rep(c(1),nrow(folds[[i]])),
                   folds[[i]][,j])
      beta <-c(fit[[i]]$mle[1],
               fit[[i]]$mle[2])
      sigma <- exp(X_1 %*% beta)
      
      nllh1.5kfold[i] <- -sum(dgpd(folds[[i]]$snow_fall,
                                   sigma = sigma,
                                   xi = fit[[i]]$mle[3],
                                   log.d = TRUE, u = 6))
    }
    fit1.5kfold[j] <- (sum(nllh1.5kfold)/5)
  }
  mylistpacked[[k]] <- fit1.5kfold
}

mylistunpacked <- list()
for (i in 1:12) {
  templist <- numeric()
  for (j in 1:50) {
    templist[j] <- mylistpacked[[j]][i]
  }
  mylistunpacked[i] <- sum(as.numeric(templist)) / 50
}

for (i in 1:12) {
  print(c(mylistunpacked[[i]],names(alps_extremes)[i]))
}

# Add second covariate, i.e. ydat = altitude
```

### Two covariates

```{r}
mylistpacked <- list()
for (k in 1:50) {
  fit2.5kfold <- list()
  folds <- shuffleshuffle()[[1]]
  foldtraining <- shuffleshuffle()[[2]]
  for (j in 1:12) {
    nllh2.5kfold <- numeric()
    fit <- list()
    if (colnames(alps_extremes)[j] != "alt") {
      for (i in 1:5) {
        fit[[i]] <- gpd.fit(xdat = foldtraining[[i]]$snow_fall,
                    ydat = cbind(foldtraining[[i]]$alt,
                                 foldtraining[[i]][,j]),
                    threshold = u, 
                    sigl = c(1,2), siglink = exp)
        X_1 <- cbind(rep(c(1),nrow(folds[[i]])),
                     folds[[i]]$alt,
                     folds[[i]][,j])
        beta <-c(fit[[i]]$mle[1],
                 fit[[i]]$mle[2],
                 fit[[i]]$mle[3])
        sigma <- exp(X_1 %*% beta)
        
        nllh2.5kfold[i] <- -sum(dgpd(folds[[i]]$snow_fall,
                                     sigma = sigma,
                                     xi = fit[[i]]$mle[4],
                                     log.d = TRUE, u = 6))
        fit2.5kfold[j] <- (sum(nllh2.5kfold)/5)
        }
      } else {
        fit2.5kfold[j] <- 0
    }
  }
  mylistpacked[[k]] <- fit2.5kfold
}

mylistunpacked <- list()
for (i in 1:12) {
  templist <- numeric()
  for (j in 1:50) {
    templist[j] <- mylistpacked[[j]][i]
  }
  mylistunpacked[i] <- sum(as.numeric(templist)) / 50
}

for (i in 1:12) {
  print(c(mylistunpacked[[i]],names(alps_extremes)[i]))
}
# Add ninth covariate, i.e. ydat = altitude + isotropy

```

### Three covariates

```{r}
mylistpacked <- list()
for (k in 1:50) {
  fit3.5kfold <- list()
  folds <- shuffleshuffle()[[1]]
  foldtraining <- shuffleshuffle()[[2]]
  for (j in 1:12) {
    nllh3.5kfold <- numeric()
    fit <- list()
      if (colnames(alps_extremes)[j] != "alt" &
          colnames(alps_extremes)[j] != "isotropy"){
      for (i in 1:5) {
        fit[[i]] <- gpd.fit(xdat = foldtraining[[i]]$snow_fall,
                        ydat = cbind(foldtraining[[i]]$alt,
                                     foldtraining[[i]]$isotropy,
                                     foldtraining[[i]][,j]),
                        threshold = u, 
                        sigl = c(1,2,3), siglink = exp)
        X_1 <- cbind(rep(c(1),nrow(folds[[i]])),
                       folds[[i]]$alt,
                       folds[[i]]$isotropy,
                       folds[[i]][,j])
        beta <-c(fit[[i]]$mle[1],
                   fit[[i]]$mle[2],
                   fit[[i]]$mle[3],
                   fit[[i]]$mle[4])
        sigma <- exp(X_1 %*% beta)
          
        nllh3.5kfold[i] <- -sum(dgpd(folds[[i]]$snow_fall,
                                       sigma = sigma,
                                       xi = fit[[i]]$mle[5],
                                       log.d = TRUE, u = 6))
        fit3.5kfold[j] <- (sum(nllh3.5kfold)/5)
        }
      } else {
        fit3.5kfold[j] <- 0
    }
  }
  mylistpacked[[k]] <- fit3.5kfold
}

mylistunpacked <- list()
for (i in 1:12) {
  templist <- numeric()
  for (j in 1:50) {
    templist[j] <- mylistpacked[[j]][i]
  }
  mylistunpacked[i] <- sum(as.numeric(templist)) / 50
}

for (i in 1:12) {
  print(c(mylistunpacked[[i]],names(alps_extremes)[i]))
}
# Add tenth covariate, i.e. ydat = altitude + isotropy + landsea
```

### Four covariates

```{r}
mylistpacked <- list()
for (k in 1:50) {
  fit4.5kfold <- list()
  folds <- shuffleshuffle()[[1]]
  foldtraining <- shuffleshuffle()[[2]]
  for (j in 1:12) {
    nllh4.5kfold <- numeric()
    fit <- list()
      if (colnames(alps_extremes)[j] != "alt" &
          colnames(alps_extremes)[j] != "isotropy" &
          colnames(alps_extremes)[j] != "land_sea"){
      for (i in 1:5) {
        fit[[i]] <- gpd.fit(xdat = foldtraining[[i]]$snow_fall,
                            ydat = cbind(foldtraining[[i]]$alt,
                                         foldtraining[[i]]$isotropy,
                                         foldtraining[[i]]$land_sea,
                                         foldtraining[[i]][,j]),
                            threshold = u,
                            sigl = c(1,2,3,4), siglink = exp)
        X_1 <- cbind(rep(c(1),nrow(folds[[i]])),
                     folds[[i]]$alt,
                     folds[[i]]$isotropy,
                     folds[[i]]$land_sea,
                     folds[[i]][,j])
        beta <-c(fit[[i]]$mle[1],
                 fit[[i]]$mle[2],
                 fit[[i]]$mle[3],
                 fit[[i]]$mle[4],
                 fit[[i]]$mle[5])
        sigma <- exp(X_1 %*% beta)
          
        nllh4.5kfold[i] <- -sum(dgpd(folds[[i]]$snow_fall,
                                       sigma = sigma,
                                       xi = fit[[i]]$mle[6],
                                       log.d = TRUE, u = 6))
        fit4.5kfold[j] <- (sum(nllh4.5kfold)/5)
        }
      } else {
        fit4.5kfold[j] <- 0
    }
  }
  mylistpacked[[k]] <- fit4.5kfold
}

mylistunpacked <- list()
for (i in 1:12) {
  templist <- numeric()
  for (j in 1:50) {
    templist[j] <- mylistpacked[[j]][i]
  }
  mylistunpacked[i] <- sum(as.numeric(templist)) / 50
}

for (i in 1:12) {
  print(c(mylistunpacked[[i]],names(alps_extremes)[i]))
}
# Add fifth covariate, i.e. ydat = altitude + isotropy + landsea + dewpoint temp
```

### Five covariates

```{r}
mylistpacked <- list()
for (k in 1:50) {
  fit5.5kfold <- list()
  folds <- shuffleshuffle()[[1]]
  foldtraining <- shuffleshuffle()[[2]]
  for (j in 1:12) {
    nllh5.5kfold <- numeric()
    fit <- list()
      if (colnames(alps_extremes)[j] != "alt" &
          colnames(alps_extremes)[j] != "isotropy" &
          colnames(alps_extremes)[j] != "land_sea" &
          colnames(alps_extremes)[j] != "dewpoint_temp"){
      for (i in 1:5) {
        fit[[i]] <- gpd.fit(xdat = foldtraining[[i]]$snow_fall,
                            ydat = cbind(foldtraining[[i]]$alt,
                                         foldtraining[[i]]$isotropy,
                                         foldtraining[[i]]$land_sea,
                                         foldtraining[[i]]$dewpoint_temp,
                                         foldtraining[[i]][,j]),
                            threshold = u,
                            sigl = c(1,2,3,4,5), siglink = exp)
        X_1 <- cbind(rep(c(1),nrow(folds[[i]])),
                     folds[[i]]$alt,
                     folds[[i]]$isotropy,
                     folds[[i]]$land_sea,
                     folds[[i]]$dewpoint_temp,
                     folds[[i]][,j])
        beta <-c(fit[[i]]$mle[1],
                 fit[[i]]$mle[2],
                 fit[[i]]$mle[3],
                 fit[[i]]$mle[4],
                 fit[[i]]$mle[5],
                 fit[[i]]$mle[6])
        sigma <- exp(X_1 %*% beta)
          
        nllh5.5kfold[i] <- -sum(dgpd(folds[[i]]$snow_fall,
                                       sigma = sigma,
                                       xi = fit[[i]]$mle[7],
                                       log.d = TRUE, u = 6))
        fit5.5kfold[j] <- (sum(nllh5.5kfold)/5)
        }
      } else {
        fit5.5kfold[j] <- 0
    }
  }
  mylistpacked[[k]] <- fit5.5kfold
}

mylistunpacked <- list()
for (i in 1:12) {
  templist <- numeric()
  for (j in 1:50) {
    templist[j] <- mylistpacked[[j]][i]
  }
  mylistunpacked[i] <- sum(as.numeric(templist)) / 50
}

for (i in 1:12) {
  print(c(mylistunpacked[[i]],names(alps_extremes)[i]))
}
# Add first covariate, i.e. ydat = altitude + isotropy + landsea + dewpoint temp + air temp
```

### Six covariates

```{r}
mylistpacked <- list()
for (k in 1:50) {
  fit6.5kfold <- list()
  folds <- shuffleshuffle()[[1]]
  foldtraining <- shuffleshuffle()[[2]]
  for (j in 1:12) {
    nllh6.5kfold <- numeric()
    fit <- list()
      if (colnames(alps_extremes)[j] != "alt" &
          colnames(alps_extremes)[j] != "isotropy" &
          colnames(alps_extremes)[j] != "land_sea" &
          colnames(alps_extremes)[j] != "dewpoint_temp" &
          colnames(alps_extremes)[j] != "air_temp"){
      for (i in 1:5) {
        fit[[i]] <- gpd.fit(xdat = foldtraining[[i]]$snow_fall,
                            ydat = cbind(foldtraining[[i]]$alt,
                                         foldtraining[[i]]$isotropy,
                                         foldtraining[[i]]$land_sea,
                                         foldtraining[[i]]$dewpoint_temp,
                                         foldtraining[[i]]$air_temp,
                                         foldtraining[[i]][,j]),
                            threshold = u,
                            sigl = c(1,2,3,4,5,6), siglink = exp)
        X_1 <- cbind(rep(c(1),nrow(folds[[i]])),
                     folds[[i]]$alt,
                     folds[[i]]$isotropy,
                     folds[[i]]$land_sea,
                     folds[[i]]$dewpoint_temp,
                     folds[[i]]$air_temp,
                     folds[[i]][,j])
        beta <-c(fit[[i]]$mle[1],
                 fit[[i]]$mle[2],
                 fit[[i]]$mle[3],
                 fit[[i]]$mle[4],
                 fit[[i]]$mle[5],
                 fit[[i]]$mle[6],
                 fit[[i]]$mle[7])
        sigma <- exp(X_1 %*% beta)
          
        nllh6.5kfold[i] <- -sum(dgpd(folds[[i]]$snow_fall,
                                       sigma = sigma,
                                       xi = fit[[i]]$mle[8],
                                       log.d = TRUE, u = 6))
        fit6.5kfold[j] <- (sum(nllh6.5kfold)/5)
        }
      } else {
        fit6.5kfold[j] <- 0
    }
  }
  mylistpacked[[k]] <- fit6.5kfold
}

mylistunpacked <- list()
for (i in 1:12) {
  templist <- numeric()
  for (j in 1:50) {
    templist[j] <- mylistpacked[[j]][i]
  }
  mylistunpacked[i] <- sum(as.numeric(templist)) / 50
}

for (i in 1:12) {
  print(c(mylistunpacked[[i]],names(alps_extremes)[i]))
}
# Add f covariate, i.e. ydat = altitude + isotropy + landsea + dewpoint temp + air temp
```

When fitting the model to the training data, the model is not valid in the validation fold. A different method would be required to do model selection; however, since the choice of covariates hasn't differed from the AIC method, we will proceed to take the full model as found earlier.

# Analysis of fitted linear model

Firstly, we plot use qq plots for model diagnostic.

```{r}
fitlin <- gpd.fit(xdat = alps_extremes$snow_fall,
                  ydat = cbind(alps_extremes$alt,
                               alps_extremes$isotropy,
                               alps_extremes$land_sea,
                               alps_extremes$dewpoint_temp,
                               alps_extremes$air_temp,
                               alps_extremes$sd,
                               alps_extremes$slope,
                               alps_extremes$geopotential,
                               alps_extremes$land_sea*alps_extremes$sd),
                threshold = u, 
                sigl = c(1,2,3,4,5,6,7,8,9), siglink = exp)
gpd.diag(fitlin)
```

We now proceed to obtain confidence intervals for our parameters.

```{r}
# Non-parametric bootstrap
fit.boot <- list()
MLE.boot <- array(dim = c(11,1000))
for (i in 1:1000) {
  alps_bootstrap <- alps_extremes[sample(1:nrow(alps_extremes), replace = TRUE),]
  
  fit.boot[[i]] <- gpd.fit(xdat = alps_bootstrap$snow_fall,
                      ydat = cbind(alps_bootstrap$alt,
                                   alps_bootstrap$isotropy,
                                   alps_bootstrap$land_sea,
                                   alps_bootstrap$dewpoint_temp,
                                   alps_bootstrap$air_temp,
                                   alps_bootstrap$sd,
                                   alps_bootstrap$slope,
                                   alps_bootstrap$geopotential,
                                   alps_bootstrap$land_sea*alps_bootstrap$sd),
                      threshold = u, 
                      sigl = c(1,2,3,4,5,6,7,8,9), siglink = exp)
  for (j in 1:11) {
    MLE.boot[j,i] <-  fit.boot[[i]]$mle[j]
  }
}

MLE.ci <- list()

for (i in 1:11) {
  low <- quantile(MLE.boot[i,],0.025)
  high <- quantile(MLE.boot[i,],0.975)
  med <- quantile(MLE.boot[i,],0.5)
  
  MLE.ci[[i]] <- c(low,high,med)
}
```

We are interested in asnwering three main questions: 1. What is the shape parameter? Is it negative? 2. Which coefficients are significant in the scale parameter? 3. What is the spacial map of the scale parameter? [Link to meaning of covariates](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-single-levels-monthly-means?tab=overview).

This is the full model (with 95% confidence intervals). $$\begin{align*}
\log\sigma &= \beta_0\\
& + \beta_{altitude} \cdot \text{altitude} \\
& + \beta_{isotropy} \cdot \text{isotropy} \\
& + \beta_{land\_sea} \cdot \text{land_sea} \\
& + \beta_{dewpoint temp} \cdot \text{dewpoint temp} \\
& + \beta_{air temp} \cdot \text{air temp} \\
& + \beta_{sd} \cdot \text{sd} \\
& + \beta_{slope} \cdot \text{slope} \\
& + \beta_{geopotential} \cdot \text{geopotential} \\
& + \beta_{interaction} \cdot \text{land_sea}*\text{sd} \\
\end{align*}$$

where

$$\begin{align*}
\beta_0 &= -0.422 && (-1.284,0.753)\\
\beta_{altitude} &= 0.129 && (0.030,0.237)\\
\beta_{isotropy} &= 0.135 && (-0.001,0.274)\\
\beta_{land\_sea} &= 0.772 && (-2.772,3.964)\\
\beta_{dewpoint\;temp} &= 1.356 && (0.229,1.891)\\
\beta_{air\;temp} &= -1.388 && (-2.045,-0.159)\\
\beta_{sd} &= -0.258 && (-1.418,0.762)\\
\beta_{slope} &= 0.255 && (-0.231,0.624)\\
\beta_{geopotential} &= -0.004 && (-0.198,0.200)\\
\beta_{interaction} &= 0.613 && (-1.98,3.966)\\
\end{align*}$$

and

$$\xi = -0.163 \;\;\;\; (-0.232,-0.092)$$

We can see that the shape parameter is negative, which suggests that the distribution has a finite upper bound. Hence, given that it snows, we can predict an upper bound for the amount of snowfall.

Furthermore, we see that the significant covariates are altitude, air temperature, and dewpoint temperature.

```{r}

alps_significant <- data.frame(alt,isotropy,land_sea,dewpoint_temp,air_temp,sd,slope,geopotential,snow_fall,lon,lat,dates)
alps_significant <- na.omit(alps_significant)

alps_significant$snow_fall <- alps_significant$snow_fall * 1000
for (i in 1:8) {
  alps_significant[,i] <- (alps_significant[,i] - mean(alps_significant[alps_significant$snow_fall > 0,i])) / sd(alps_significant[alps_significant$snow_fall > 0,i])
}

date <- "2021-12"

alps_significant <- alps_significant[alps_significant$dates == date,]


X1 <- cbind(rep(c(1),nrow(alps_significant)),
                alps_significant$alt,
                alps_significant$isotropy,
                alps_significant$land_sea,
                alps_significant$dewpoint_temp,
                alps_significant$air_temp,
                alps_significant$sd,
                alps_significant$slope,
                alps_significant$geopotential,
                alps_significant$land_sea*alps_significant$sd)
                
beta.low <- numeric()
for (i in 1:10) {
  beta.low[i] <- MLE.ci[[i]][1]
}

beta.high <- numeric()
for (i in 1:10) {
  beta.high[i] <- MLE.ci[[i]][2]
}

beta.avg <- numeric()
for (i in 1:10) {
  beta.avg[i] <- MLE.ci[[i]][3]
}

sigma.low <- exp(X1 %*% beta.low)
sigma.high <- exp(X1 %*% beta.high)
sigma.avg <- exp(X1 %*% beta.avg)

returnlvl <- u - sigma.avg / MLE.ci[[11]][3]



df <- data.frame("upper_snow_mm"=c(returnlvl),
                 "lon"=c(alps_significant$lon),
                 "lat"=c(alps_significant$lat)
)


library(ggplot2)
library(ggmap)

indbox = make_bbox(lon = c(min(df$lon), max(df$lon))+c(-3,3), lat = c(max(df$lat), min(df$lat)), f = 0)
world <- map_data("world")
p = ggplot() +
  geom_map(
    data = world, map = world,
    aes(long, lat, map_id = region),
    color = "black", fill = "lightgray", size = 0.1
  ) + scale_x_continuous(limits=indbox[c(1,3)]) + scale_y_continuous(limits=indbox[c(2,4)]) + coord_fixed(ratio=1.4) +
geom_tile(df, mapping = aes(x = lon, y = lat, fill = upper_snow_mm), width = 0.27, height = 0.27) 
p

```
